---
title: "Does Race type affects a resident's life"
author: "Hu Xiangkun"
date: "2014年10月20日"
output: html_document
---

I will study the problem that whether a resident's race affects his/her life. In this project, I use the data from General Social Survey (GSS), which conducted between 1972 - 2012 from the residents of United States. The dataset names "gss". I will figure this out by contrasting the education and income level between the white and black race.

First let's load the dataset as below:

```{r}
#The gss data can also be loaded from "http://d396qusza40orc.cloudfront.net/statistics/project/gss.Rdata"
load("gss.Rdata")
```

The gss dataset has 114 variables, there are 3 variable I care about here are "race", "incom16" and "degree", so let's get the subset:

```{r}
#This project only care about race, income and eduction level
gss_p = na.omit(gss[, c("race", "incom16", "degree")])
```

Let's see the summary of this subset:

```{r}
summary(gss_p)
```

The summary infomation shows that the main race in the USA are "White" and "Black", and there are 6 types of income level and 5 types of eduction level. So here I only concern "White" and "Black" race for this study.

```{r}
#Only concern about "White" and "Black" race
data = subset(gss_p, gss_p$race == "White" | gss_p$race == "Black")
summary(data)
```

So next I will use hypothsis test and confidence interval to test whether there is a difference between White and Black in income and eduction level.

Income
===========
There are 6 income levels, I will study them one by one. As the method in study different types are the same, so I will use the income level "Far Below Average" as an example.

#### Far Below Average
Using **pw** and **pb** to represent the probabilty of the resident has a income far below the average. Cause I will use hypothsis test, now I state the hypothsis below:

**H0: Race has no effect in income level far below average, the two probabilties are equal, i.e. pw = pb** 

**H1: Race type does has an effect, i.e. pw ≠ pb**

Now let's calculate the point estimate of **pw-pb**:

```{r}

```







